Grappa: Scalable Computing on Mass-Market Hardware for Irregular, Low-Locality Applications

### 2012-07-23 doc/asplos13/tasks.tex
Revised section Tasks

The basic unit of execution in Grappa is a {\em task}. Tasks are
small; just a function pointer and arguments. Tasks are only
queued when they are spawned; later, when resources are free, they are
allocated a stack, bound to a core, and executed.

Grappa's tasks give up control of their core when when they perform
long-latency operations, allowing the processor to remain busy while
waiting for the operation to complete. This is most often done
implicitly inside a call into Grappa's API, but it can also be done
explicitly by the programmer using the calls shown in
Figure~\ref{fig:scheduling}. \TODO{cite UPC split phase reads and
writes, but include that we provide a way to overlap computation}

All Grappa scheduling is done in userspace, and minimal state is saved,
in order to minimize the context switch time. In our experiments, we saw
context switch times as low as \checkme{40ns}. This lightweight context
switching is Grappa's key enabling feature. It gives us the ability to
tolerate latency. Given that ability, we are able to \textbf{trade
latency for throughput}: by {\em increasing} latency in key components
in the system we are able to increase our aggregate random access
bandwidth, our synchronization bandwidth, and our ability to tolerate
load.

\begin{figure}[htbp]
  \begin{center}
    \begin{description}\small
    \item[ \texttt{ yield() } ] \hfill \\
      Gives up control of core to scheduler, queuing task to be scheduled again soon
    \item[ \texttt{ suspend() } ] \hfill \\
      Gives up control of core to scheduler
    \item[ \texttt{ wake( task * $t$ ) } ] \hfill \\
      Enqueues $t$ to be scheduled again soon
    \end{description}
    \begin{minipage}{0.95\columnwidth}
      \caption{\label{fig:scheduling} Grappa API: scheduling} %{-4ex}}
    \end{minipage}
    %\vspace{-3ex}
  \end{center}
\end{figure}

>>>>

The basic unit of execution in Grappa is a {\em task}. Each task is
represented by a function pointer and its arguments and is enqueued,
but not instantiated, when spawned.  Later, when resources are free,
the task is allocated a stack, bound to a core, and executed.

During execution, a task yields control of its core whenever it performs
a long-latency operation, allowing the processor to remain busy while
waiting for the operation to complete.  In addition,  a programmer can
direct scheduling explicitly via the Grappa API calls shown
in Figure~\ref{fig:scheduling}.
\TODO{cite UPC split phase reads and
writes, but include that we provide a way to overlap computation}

To minimize yield overhead, all Grappa scheduling is done in
user-space and saves minimal state, leading to context switch times as
low as \checkme{40ns} even when switching amongst thousands of tasks.
This fast context switching amongst many tasks is crucial to Grappa's
effectiveness for two reasons.  First, latency can be tolerated only
when the net cost of context switching is dwarfed by the loss incurred
by the alternative: stalling.  Second, because we can switch amongst
so many tasks efficiently, we can \textbf{trade latency for throughput}:
by {\em increasing} latency in key components of the
system we are able to increase our aggregate random access bandwidth,
our synchronization bandwidth, and our ability to tolerate load
imbalance.
\TODO{check to see we actually explain these three somewhere.  Ie, that aggregating increases throughput and latency, that delegating synchronization increases the rate and the latency at which we can eg atomically increment, and that stealing work increases the latency of some tasks (eg when their data is largely at the originating node) but provides greater throughput overall.}

\begin{figure}[htbp]
  \begin{center}
    \begin{description}\small
    \item[ \texttt{ yield() } ] \hfill \\
      Yields core to scheduler, enqueuing caller to be scheduled again soon
    \item[ \texttt{ suspend() } ] \hfill \\
      Yields core to scheduler, enqueuing caller only once another task calls wake
    \item[ \texttt{ wake( task * $t$ ) } ] \hfill \\
      Enqueues some other task $t$ to be scheduled again soon
    \end{description}
    \begin{minipage}{0.95\columnwidth}
      \caption{\label{fig:scheduling} Grappa API: scheduling} %{-4ex}}
    \end{minipage}
    %\vspace{-3ex}
  \end{center}
\end{figure}
